Domain-Driven Generative Data Model for BigPetStore

### 2014-07-16 bdcloud_paper/latex/paper.tex
Edited introduction.  Added placeholders for references

\section{Introduction}
Big Data applications process large, continuously-changing data sets and with the general goal of information and knowledge extraction.  Novel tools such as Pig, Spark, and Crunch based around Big Data frameworks are being developed every day.  These tools are often developed on desktops but need to be tested on clusters against large input sets.  The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent that can be used across a range of environments.

The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems. BigPetStore has been incorporated into the Apache BigTop distribution for testing and as a reference application.

In this work, we present a novel domain-drive model for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space and incorporate it into the BigPetStore application.   Our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter. Combining the features of the TeraGen (scalability) and MovieLens (semantically rich) input data sets commonly used for Big Data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with the same desirable properties (global patterns and region specific transactions) without some of the drawbacks (the need for hierarchical imperative coding of attribute selection during data set generation).

To benchmark, one must use inputs which are reliable, but also, which challenge a system in ways which approximate real-world usage.  Over time, more high-quality datasets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs for benchmarking. Few data sets are scalable, allowing the user to pick how much data they want.  As a result, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store. Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.  Several companies (such as Twitter) offer access to their data but the costs and usage restrictions can be undesirable.

Models for generating synthetic data overcome these obstacles.  The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.  TeraGen can randomly generate numerical data at any scale, but the data is semantically void and has very little meaning.  Thus, TeraGen cannot be used to test any application other than a simplistic “sort” program.  The Intel Hadoop Benchmark suite include several generative data models such as data for computing the PageRank of web sites for their benchmarks. The models are specific to each benchmark, however, and not useful for general-purpose development and testing.

Given that the driving motivation behind so much of the investment in big data software solutions is “to derive insight from data”, there is thus an obvious, gaping hole in the existing ecosystem: We need a tool that generates meaningful data which contains informational complexity that can be used to derive global patterns in a data set, which can be used to scalably and reliably generate large data sets on the fly - so that we can use it for benchmarking, comparison, and demonstration of the different algorithmic approaches to mining large data sets.

>>>>

\section{Introduction}
Big data applications process large, continuously-changing data sets with the general goal of information and knowledge extraction.  New tools such as Hive, Pig and Crunch \textcolor{red}{reference} based around big data frameworks such as Hadoop, Storm, and Spark \textcolor{red}{reference} are being released every day. Development, testing and benchmarking of big data software, algorithms, and infrastructure pose unique challenges.  Though software and algorithms will be deployed to large clusters for benchmarking and production usage, engineers develop and test these tools on desktops due to the speed and ease of development and debugging on a local machine and limited access to large clusters due to the high costs of purchasing, provisioning, and maintaining such clusters.  Software and algorithms are being developed at incredible speed with new releases everyday. The high speed of development driven by the need to be the first to market in a competitive field often results in buggy software.  There is great need for consistent and rigorous testing and benchmarking frameworks to aid in the identification of bugs while maintaining the high-speed of development.  Access to high-quality, semantically-rich data sets and complex, realistic analytics workflow examples that can scale from local development environments to large clusters are key to increasing access to and development of consistent and rigorour testing and benchmarking frameworks.

Over time, more high-quality data sets are becoming publicly available, but these data sets often carry burdens that force developers to make undesirable trade offs. Since few data sets are scalable, allowing the user to pick how much data they want, users are forced to choose between small data sets which are useful for development and testing on a local machine, but not for production-scale tests or benchmarking, or large data sets which are prohibitively expensive to host, transfer, and store. Furthermore, access to large data sets can be unreliable due to dependence on third party benefactors for hosting and requires significant amounts of bandwidth to download.  Several companies (such as Twitter \textcolor{red}{reference}) offer access to their data but the costs and usage restrictions can be undesirable.

\emph{Ab initio} mathematical models for generating synthetic data overcomes several of these obstacles.  The code for generating the data is often small, enabling free hosting, licensing restrictions are often avoided, and the data size can be scaled, enabling seamless usage of the same data set across a range of scenarios.  Common synthetic data set generators such as TeraGen and the Intel Hadoop Benchmark \textcolor{red}{reference} can generate data sets quickly and at any scale, but they are over-specialized for benchmarking, resulting in semantically-void data that lacks usefulness for analytics examples. The lack of frameworks for generating semantically-rich data at arbitrary scale denies developers access to consistent data sets that can be used across a range of environments \textcolor{red}{Huppler, Pavlo}.

Attempts have been made to find a middle-ground.  Myriad and PDGF provide primitives for specifying generative data models \textcolor{red}{reference}. Alexandrov, et al. proposed Oligos, an approach for generating synthetic data from models trained on reference databases with real data \textcolor{red}{reference}. Oligos appears promising but will need to overcome several difficult challenges in automatically characterizing complex reference data to determine relevant constraints and learn accurate models.  Until such methods have reached maturity, development of a customized domain-driven model is best suited for generating for satisfying the needs of big data software and algorithms development.

The BigPetStore application aims to be a big data application blueprint built around a fictional chain of pet stores, which incorporates a data generator and complex examples of cleaning, aggregation, and analytics using well-known tools such as Hive and Mahout from across the Hadoop and Spark ecosystems \textcolor{red}{reference}. BigPetStore's usefulness has been demonstrated by its incorporation into the Apache BigTop distribution for testing and as a reference application \textcolor{red}{reference}.

In this work, we present a new domain-drive model and simulation for generating arbitrarily large, complex data for testing and benchmarking of software tools in the big data space. Compared with BigPetStore's previous data generator, our model and implementation can generate data which contains geospatial, temporal, and quantitative features, similar to the type of data which businesses and organizations might typically encounter. Combining the features of the TeraGen (scalability) \textcolor{red}{reference} and MovieLens (semantically rich) \textcolor{red}{reference} input data sets commonly used for big data benchmarking, we present a scalable synthetic data set generation implementation which can be used to benchmark lower level tasks (such as sorting) as well as higher level tasks (such as clustering, recommending, and search indexing) at arbitrary scale.  To validate this approach, we apply it as a drop in replacement for the current version of BigPetStore which has already been utilized to test several Hadoop deployments at terabyte scale, and is actively being adopted by a variety of companies in the software industry, and confirm that it generates data with desirable properties (global patterns and region specific transactions).

%Models for synthetic data generation is not the only approach for obtaining data sets. 
%can randomly generate numerical data at any scale, but the data is semantically void and has very little meaning.  Thus, TeraGen cannot be used to test any application other than a simplistic sort program.  The Intel Hadoop Benchmark suite include several generative data models such as data for computing the PageRank of web sites for their benchmarks. The models are specific to each benchmark, however, and not useful for general-purpose development and testing.

%Given that the driving motivation behind so much of the investment in big data software solutions is ``to derive insight from data,'' there is thus an obvious, gaping hole in the existing ecosystem: We need a tool that generates meaningful data which contains informational complexity that can be used to derive global patterns in a data set, which can be used to scalably and reliably generate large data sets on the fly - so that we can use it for benchmarking, comparison, and demonstration of the different algorithmic approaches to mining large data sets.
